Synthesizing Symmetric Lenses

### 2018-06-11 papers/SymmetricLenses/main.tex
remove rights stuff
Added comments to Sections 3 and 4

\setcopyright{rightsretained}
\acmPrice{}
\acmDOI{10.1145/3158089}
\acmYear{2019}
\copyrightyear{2019}
\acmJournal{PACMPL}
\acmVolume{3}
\acmNumber{POPL}
\acmArticle{1}
\acmMonth{1}
\startPage{1}

@@

% begin ssl
\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. These lens terms are typed between pairs of regular expressions: the
typing judgment $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ means that
$\Lens$ is a well typed lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$.

@@

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting is particularly useful when chaining many compositions together, as it
can be used to align the central types.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \equiv \Regex'\\
    \RegexAlt \equiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }
\]
Type equivalence is useful both for addressing type annotations and for
letting compositions go through.
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting is particularly useful when chaining many compositions together, as it
can be used to align the central types.
% end ssl

@@

The language is defined like the regular expression without probability
annotations would be.  The probabilities are defined as follows:

\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

@@

    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{\Character_1 \ldots \Character_n}$

@@

Stochastic DNF regular expressions are hierarchically built, where stochastic
DNF regular expressions are lists of stochastic sequences. Stochastic sequences
themselves are lists of interleaved strings and stochastic atoms.  Stochastic
atoms are iterated stochastic DNF regular expressions.
\begin{center}
  \begin{tabular}{l@{\ }c@{\ }l@{\ }>{\itshape\/}r}
    % DNF_REGEX
    \Atom{},\AtomAlt{} & \GEq{} & \PRegexStar{\DNFRegex{}}{\Probability}
% & \StarAtomType{}
\\
    \Sequence{},\SequenceAlt{} & \GEq{} &
                                                       $\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}$ 
%& \MultiConcatSequenceType{} 
\\
    \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$ %& \MultiOrDNFRegexType{} 
  \end{tabular}
\end{center}

Intuitively, stochastic DNF regular expressions are essentially stochastic
regular expressions with all concatenations fully distributed over all
disjunctions. As such, the language of a stochastic DNF regular expression is a
union of its subcomponents, the language of a stochastic sequence is the
concatenation of its subcomponents, and the language of a stochastoc atom is the
iteration of its subcomponent.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic. If $\Regex$ is a
stochastic DNF regular expression generated by \ToStochastic, then
$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequences, and gives
iterations a .8 chance to continue iterating.

\paragraph*{Entropy} In the situation where a DNF regular expression is
unambiguous, we have defined an efficient, syntactic means to find the entropy
of the DNF regular expression.

\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
    & =
    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
      - \Log_2(1-\Probability)
      $\\

@@

In greater detail, a symmetric DNF Lens \SDNFLens{} consists of three
components. First is the mapping component. The mapping component is a list of
triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component, \SSQLens{}, is a sequence lens that maps between the two
sequences specified by $i$ and $j$. The second component is a list of integers
$c$. The integer, $c_i$ at position $i$ states that when performing a $\CreateR$
on a string matching sequence $\Sequence_i$, it will use the sequence lens that
maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third component is a
list of integers $d$, and follows the same pattern as the second component, but
maps from right to left.

A Sequence lens \SSQLens{} consists of three components. First is the mapping
component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
sequences, respectively. The third component, \SAtomLens{}, is an atom lens that
maps between the two atoms specified by $i$ and $j$. The second component is a
list of strings, $\String$. The string, $s_i$ at position $i$ states that when
performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
two things. If $\Atom_i$ is mapped to in the mapping component, then performing
the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
not mapped to, then $\String_i$ will be created as a representative of
$\Atom_i$.

@@

\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.

@@

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
  & = 
  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
    \StringAlt_r'' \Concat \StringAlt_r'$\\
  & & where $\StringAlt_y'' =
    \begin{cases*}
      \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_y} & if $j_k = y$\\
      \StringAlt_y & if $\nexists k. j_k = y$\\
    \end{cases*}$\\
  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
  \Concat \StringAlt_q'$
  & = 
  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
    \String_r'' \Concat \String_r'$\\
  & & where $\String_x'' =
    \begin{cases*}
      \SAtomLens_k.\PutROf{\StringAlt_{j_k}''}{\String_x} & if $i_k = x$\\
      \String_x & if $\nexists k. i_k = x$\\
    \end{cases*}$\\
  $\PutR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q' \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q'$
  & =
  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
  & & where
      $t_y''' =
      \begin{cases*}
        \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_{y}''} & if $j_k = y$\\
        \StringAlt_y'' & if $\nexists k. j_k = y$\\
      \end{cases*}$\\
  $\PutL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
  & =
  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
  & & where
      $s_x''' =
      \begin{cases*}
        \SAtomLens_k.\PutLOf{\StringAlt_{j_k}''}{\String_{x}''} & if $i_k = x$\\
        \String_x'' & if $\nexists k. i_k = x$\\
      \end{cases*}$\\
\end{tabular}

@@

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
  & = 
  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
  & = 
  & $\String_1 \Concat \ldots \Concat \String_m$
  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
  $\PutR{} \App \String_0 \Concat \ldots \Concat \String_n \App
  \StringAlt_1 \Concat \ldots \Concat \StringAlt_m$
  & = 
  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
    where $\StringAlt_i' =
    \begin{cases*}
      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
      \SDNFLens.\CreateR \App \String_i & otherwise
    \end{cases*}$\\
  $\PutL{} \App \StringAlt_1 \Concat \ldots \Concat \StringAlt_m \App
  \String_0 \Concat \ldots \Concat \String_n$
  & = 
  & $\String_1' \Concat \ldots \Concat \String_m'$
    where $\String_i' =
    \begin{cases*}
      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
    \end{cases*}$\\
\end{tabular}

@@

Instead of searching for \emph{any} term that satisfies the specification, we
search for the \emph{best} term that satisfies the specification. This begs the
question, by what metric do we use to determine which term is best. Intuitively,
we aim to synthesize the lens which is \emph{closest} to being a bijection.
Somewhat more formally, we aim to synthesize a lens which minimizes the expected
number of bits which must be used to recover the source from the target, and the
target from the source. The expected number of bits that are needed to encode
informatation from a data source is a well known concept in information theory,
and is known as the \emph{entropy} of the data source.

The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
  \Given \DNFRegex,\SDNFLens}$.

\begin{centering}
  \[
    \begin{array}{rl}
      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
      & \hspace*{1em}\Given
        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
        \ldots\DNFLSep
        (i_p,j_p,\SSQLens_p)}
        ,\ListOf{c_1;\ldots;c_q}
        ,\ListOf{d_1;\ldots;d_r})}\\
      =\\
      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
      \\

@@

The greedy algorithm is actually two greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth).

\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SSQLens) \gets
    \Call{GreedyState.PopBest}{\GreedyState}$
    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
    \EndIf
    \EndWhile
    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
input-output examples. First, \PCF{CannotMap} checks if there is possibly a lens
between the given DNF regular expressions that is consistent with the examples.
If there is not, \GreedySynth returns \None early. After this, \GreedySynth
finds the best lenses between each sequence pair of the left and right DNF
regular expressions. An internal state for the algorithm is initialized from
these. This state consists of two things: an intermediate partial DNF Lens and a
priority queue consisting of potential sequence lenses. On initialization,
\GreedySynth checks which sequence lenses are \emph{required} to be included,
from the examples. Those are added into the partial lens, and the rest are
thrown into the priority queue.

>>>>

  \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

%\setcopyright{rightsretained}
%\acmPrice{}
%\acmDOI{10.1145/3158089}
%\acmYear{2019}
%\copyrightyear{2019}
%\acmJournal{PACMPL}
%\acmVolume{3}
%\acmNumber{POPL}
%\acmArticle{1}
%\acmMonth{1}
%\startPage{1}

@@

% begin ssl
\section{Simple Symmetric String Lenses}
\label{sec:ssl}
\sam{Perhaps we shoud introduce stochastic regular expressions first before introducing the lens combinators? This is because the type equivalence combinator mentions regexp equivalence, but this has not been introduced by this point in the paper.}
We now give a formal description of all the combinators in our symmetric lens
language. These lens terms are typed between pairs of regular expressions: the
typing judgment $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ means that
$\Lens$ is a well typed lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$.

@@

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa. The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \equiv \Regex'\\
    \RegexAlt \equiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }
\]
Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is equivalent to $T'$. Type equivalence is useful both for addressing type annotations and for
letting compositions go through.\sam{I do not understand this statement.}
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
% end ssl

@@

The language is defined like the regular expression without probability
annotations would be.  The probabilities are defined as follows:
\sam{Why in the definition of star do we multiply by $p^n * (1-p)$? I think this 
should be explained}
\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

@@

    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$

@@

Stochastic DNF regular expressions are hierarchically built, where stochastic
DNF regular expressions are lists of stochastic sequences. Stochastic sequences
themselves are lists of interleaved strings and stochastic atoms.  Stochastic
atoms are iterated stochastic DNF regular expressions.
\sam{I think we should explicitly say what sets $A, B$, $SQ, TQ$ and $DS, DT$ range over.}
\begin{center}
  \begin{tabular}{l@{\ }c@{\ }l@{\ }>{\itshape\/}r}
    % DNF_REGEX
    \Atom{},\AtomAlt{} & \GEq{} & \PRegexStar{\DNFRegex{}}{\Probability}
% & \StarAtomType{}
\\
    \Sequence{},\SequenceAlt{} & \GEq{} &
                                                       $\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}$ 
%& \MultiConcatSequenceType{} 
\\
    \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$ %& \MultiOrDNFRegexType{} 
  \end{tabular}
\end{center}

Intuitively, stochastic DNF regular expressions are essentially stochastic
regular expressions with all concatenations fully distributed over all
disjunctions. As such, the language of a stochastic DNF regular expression is a
union of its subcomponents, the language of a stochastic sequence is the
concatenation of its subcomponents, and the language of a stochastic atom is the
iteration of its subcomponent.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic. If $\Regex$ is a
stochastic DNF regular expression generated by \ToStochastic, then
$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequences, and gives
iterations a .8 chance to continue iterating.\sam{I think we should mention that
.8 is chosen because it works well in practice.}

\paragraph*{Entropy}
\sam{We should define what entropy means and why it is significant}
In the situation where a DNF regular expression is
unambiguous, we have defined an efficient, syntactic means to find the entropy
of the DNF regular expression.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
    & =
    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
      - \Log_2(1-\Probability)
      $\\

@@

In greater detail, a symmetric DNF Lens \SDNFLens{} consists of three
components. First is the mapping component. The mapping component is a list of
triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between the two
sequences specified by $i$ and $j$. The second component is a list of integers
$c$. The integer, $c_i$ at position $i$ states that when performing a $\CreateR$
on a string matching sequence $\Sequence_i$, it will use the sequence lens that
maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third component is a
list of integers $d$, and follows the same pattern as the second component, but
maps from right to left.

A Sequence lens \SSQLens{} consists of three components. First is the mapping
component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
sequences, respectively. The third component of the triple, \SAtomLens{}, is an atom lens that
maps between the two atoms specified by $i$ and $j$. The second component is a
list of strings, $\String$. The string, $s_i$ at position $i$ states that when
performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
two things. If $\Atom_i$ is mapped to in the mapping component, then performing
the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
not mapped to, then $\String_i$ will be created as a representative of
$\Atom_i$.

@@

\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
\sam{I found that I understood the semantics easier than the typing
rules. Perhaps we should switch the order that we do them in?}
The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.

@@

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
  & = 
  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
    \StringAlt_r'' \Concat \StringAlt_r'$\\
  & & where $\StringAlt_y'' =
    \begin{cases*}
      \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_y} & if $j_k = y$\\
      \StringAlt_y & if $\nexists k. j_k = y$\\
    \end{cases*}$\\
  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
  \Concat \StringAlt_q'$
  & = 
  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
    \String_r'' \Concat \String_r'$\\
  & & where $\String_x'' =
    \begin{cases*}
      \SAtomLens_k.\PutROf{\StringAlt_{j_k}''}{\String_x} & if $i_k = x$\\
      \String_x & if $\nexists k. i_k = x$\\
    \end{cases*}$\\
  $\PutR{} \App (\String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q') \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q')$
  & =
  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
  & & where
      $t_y''' =
      \begin{cases*}
        \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_{y}''} & if $j_k = y$\\
        \StringAlt_y'' & if $\nexists k. j_k = y$\\
      \end{cases*}$\\
  $\PutL{} \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'') \Concat (\StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q')$
  & =
  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
  & & where
      $s_x''' =
      \begin{cases*}
        \SAtomLens_k.\PutLOf{\StringAlt_{j_k}''}{\String_{x}''} & if $i_k = x$\\
        \String_x'' & if $\nexists k. i_k = x$\\
      \end{cases*}$\\
\end{tabular}

@@

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
  & = 
  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
  & = 
  & $\String_1 \Concat \ldots \Concat \String_m$
  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
  $\PutR{} \App (\String_0 \Concat \ldots \Concat \String_n) \App
  (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m)$
  & = 
  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
    where $\StringAlt_i' =
    \begin{cases*}
      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
      \SDNFLens.\CreateR \App \String_i & otherwise
    \end{cases*}$\\
  $\PutL{} \App (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m) \App
  (\String_0 \Concat \ldots \Concat \String_n)$
  & = 
  & $\String_1' \Concat \ldots \Concat \String_m'$
    where $\String_i' =
    \begin{cases*}
      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
    \end{cases*}$\\
\end{tabular}

@@

Instead of searching for \emph{any} term that satisfies the specification, we
search for the \emph{best} term that satisfies the specification. This begs the
question: what metric should we use to determine which term is best? Intuitively,
we aim to synthesize the lens which is \emph{closest} to being a bijection.
Somewhat more formally, we aim to synthesize a lens which minimizes the expected
number of bits which must be used to recover the source from the target, and the
target from the source. The expected number of bits that are needed to encode
informatation from a data source is a well known concept in information theory,
and is known as the \emph{entropy} of the data source.

The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
  \Given \DNFRegex,\SDNFLens}$.
\sam{We should definitely explain the formulae for entropy computing entropy. Also I did not understand $argmin_{\theta}$ terms were being talked about in the 
previous sentence.}
\begin{centering}
  \[
    \begin{array}{rl}
      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
      & \hspace*{1em}\Given
        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
        \ldots\DNFLSep
        (i_p,j_p,\SSQLens_p)}
        ,\ListOf{c_1;\ldots;c_q}
        ,\ListOf{d_1;\ldots;d_r})}\\
      =\\
      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
      \\

@@

The greedy algorithm is actually two greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth).
\sam{Could it be better to describe the algorithm in English first before giving
pseudo code? I think that putting the algorithm first might cause readers to read the 
pseudo code and get confused whereas if the algorithm is first described in English then they have a better chance of understanding it.}
\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SSQLens) \gets
    \Call{GreedyState.PopBest}{\GreedyState}$
    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
    \EndIf
    \EndWhile
    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}
